---
title: "A realistic aggregate study: estimation, dichotomization, and tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A realistic aggregate study: estimation, dichotomization, and tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(1)
```

This vignette analyzes a realistic (but homogeneous-preference) study: two
3-level attributes plus price, four alternatives per task, a five-point
purchase-likelihood follow-up, and 3,600 tasks (think 300 respondents times
12 tasks).

## Simulate and estimate

```{r}
library(ordinalDR)
beta <- c(a2 = 0.8, a3 = -0.5, b2 = 0.4, b3 = 1.0, price = -0.9)
cut <- c(-1.0, 0.2, 1.2, 2.2)
des <- dr_design(3600, J = 4, seed = 42)
dat <- dr_simulate(des, matrix(beta, 1, 5), cut, model = "B", seed = 43)
table(dat$y) / length(dat$y)
fit <- dr_mle(dat)
fit
```

## What dichotomizing the scale costs

Collapsing the five-point response at the top-2-box turns the data into the
binary dual response of the literature. The binary model is *correctly
specified* (collapsing a cumulative-link model at one of its own cut points
is exact), so the cost is efficiency, not bias:

```{r}
dat_bin <- dr_dichotomize(dat, k = 4)
fit_bin <- dr_mle(dat_bin)
round(rbind(ordinal = fit$se_beta, binary_t2b = fit_bin$se_beta), 4)
```

The Fisher machinery quantifies this exactly. The second stage contributes
information weight $\omega$ per task; dichotomization discards part of it:

```{r}
mu <- log(rowSums(exp(matrix(des$X %*% beta, ncol = 4, byrow = TRUE))))
c(W5 = mean(dr_omega(mu, cut, "B")),
  top2box = mean(dr_omega(mu, cut[3], "B")),
  topbox = mean(dr_omega(mu, cut[4], "B")))
```

## Specification tests

Two one-degree-of-freedom tests come with the model: the unit slope on the
inclusive value ($\lambda = 1$), and the conditioning parameter that spans
our model ($\delta = 0$) and DR-2Max ($\delta = 1$):

```{r}
dr_spec_tests(dat, mle = fit)
```

On data generated by the model, $\lambda = 1$ and $\delta = 0$ are retained
and DR-2Max is decisively rejected.

## Designing the scale

How many scale points, and where should the cut points sit? The
information-optimal placement and the equal-share placement for a typical
task mix:

```{r}
dr_optimal_cuts(5, "B", m0 = 0.6, s = 0.5)
dr_equal_share_cuts(5, "B", m0 = 0.6, s = 0.5)
```
