---
title: "The model in miniature: a first ordinal dual-response example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The model in miniature: a first ordinal dual-response example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette walks through the smallest interesting version of the ordinal
dual-response model: **one attribute**, homogeneous preferences, and a
**three-point** purchase-likelihood scale ("unlikely / unsure / likely").
It mirrors the worked example in the accompanying paper.

## The behavioral model in four lines

In each task a consumer sees $J$ inside goods with utilities
$u_j = V_j + \eta_j$, $\eta_j \sim \text{Gumbel}(0,1)$, picks the best one
(the forced choice), and then evaluates buying it against an outside good
whose shock $\eta_0$ she has not resolved. Under **Model A** she reports the
*probability* $F(u^*)$ in bins; under **Model B** she grades the realized
comparison $u^* - \eta_0$. Both give a joint likelihood of MNL $\times$ a
cumulative-link ordinal model driven by the task's inclusive value
$\bar\mu = \ln \sum_j e^{V_j}$.

## Two consumers, opposite reports, identical choices

Give two consumers the same taste ($\beta = 1$) and the same *within-task*
spread of the attribute, but different levels:

```{r}
library(ordinalDR)
beta <- 1
cut <- c(0, 2)                       # W = 3: "unlikely" | "unsure" | "likely"
x1 <- c(2, 1)                        # attractive task
x2 <- c(-1, -2)                      # unattractive task

mnl <- function(x) exp(beta * x) / sum(exp(beta * x))
rbind(task1 = mnl(x1), task2 = mnl(x2))
```

The forced-choice probabilities are identical — the MNL reads only
*differences*. The ordinal stage reads *levels*:

```{r}
mubar <- c(log(sum(exp(beta * x1))), log(sum(exp(beta * x2))))
probs <- function(mu, model) diff(c(0, if (model == "B") plogis(cut - mu)
                                    else exp(-exp(-(cut - mu))), 1))
rbind(task1_B = probs(mubar[1], "B"), task2_B = probs(mubar[2], "B"),
      task1_A = probs(mubar[1], "A"), task2_A = probs(mubar[2], "A"))
```

Near mirror images. All demand information flows through the second response.

## Simulate the two-consumer world at scale and estimate

`dr_simulate()` draws raw utilities and applies the reporting rules — the
estimation below therefore validates the *derivation*, not just the code.

```{r}
n <- 4000
X <- matrix(rep(c(x1, x2), n / 2), ncol = 1)
des <- list(X = X, n_tasks = n, J = 2, P = 1)
dat <- dr_simulate(des, matrix(beta, 1, 1), cut, model = "B", seed = 7)
fit <- dr_mle(dat)
fit
```

The MLE recovers $\beta = 1$ and the cut points $(0, 2)$ from nothing but
choices and three-point reports. The identification requirement — no
attribute direction constant across all inside goods — holds here because
the two tasks sit at different levels:

```{r}
dr_rank_check(des$X)
```

## Where to go next

* `vignette("aggregate-study")` — a realistic aggregate analysis with
  dichotomization comparisons and specification tests.
* `vignette("hierarchical")` — the hierarchical Bayesian model with
  individual part-worths and heterogeneous cut points.
